---
title: "hw4_bdevapa"
format: html
---

# Task 1

## Question 1 - function rewrite

```{r}
#1
library(dplyr)

iris_tbl <- as_tibble(iris)
iris_tbl |>
  filter(Petal.Length < 1.55) |>
  arrange(Species) |>
  select(starts_with("Petal"), Species)
```

## Questions 2 - Long Format vs Wide Format Data and which one is preferred

Long Format: Each of the observation is a record Wide Format: Multiple observations for the same subject maybe spread across multiple columns. Long Format is generally preferred for analysis and visualization per lecture notes

# Task 2 -Chaining where possible

## Question 3 & 4 - Glass Data

```{r}

library(dplyr)

# glass_data is already a tibble 

glass_data_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                         delim = ",",
                         col_names = c("Id","RI","Na","Mg","Al","Si","K","Ca","Ba","Fe","Type of Glass"))

glass_data_tbl |>
  mutate (Type_of_glass = factor(ifelse(`Type of Glass` == 1, "building_windows_float_processed", 
                                         ifelse(`Type of Glass` == 2, "building_windows_non_float_processed",
                                                ifelse(`Type of Glass` == 3, "vehicle_windows_float_processed",
                                                       ifelse(`Type of Glass` == 4, "vehicle_windows_non_float_processed",
                                                              ifelse(`Type of Glass` == 5, "containers",
                                                                     ifelse(`Type of Glass` == 6, "tableware",
                                                                          ifelse(`Type of Glass` == 7, "headlamps",  "unknown"
                                                                     ))))))))) |>
  filter(Fe < 0.2, Type_of_glass %in% c("tableware", "headlamps")) |>
  select(everything())

```

## Question 5 & 6 - Yeast Data

```{r}

yeast_data_tbl <- read_fwf("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         fwf_widths (c(12,6,6,6,6,6,6,6,6,3),
                         c("seq_name","mcg","gvh","alm","mit","erl","pox","vac","nuc","class")))

yeast_data_tbl |>
  select(-c(seq_name,nuc)) |>
  group_by(class) |>
  mutate( #keeps original columns 
    across(
      everything(),
      list(mean = mean, median = median)
    )
  ) |>
  select(everything())

```

## Question 7, 8, 9 & 10 - Wine Data

```{r}
library(readxl)
library(readr)

white_wine_data <- read_excel("data/white-wine.xlsx")

col_names_data <- read_excel("data/white-wine.xlsx", 
                            sheet = excel_sheets("data/white-wine.xlsx")[2])
colnames(white_wine_data) <- col_names_data$Variables
white_wine_data$type <- c("white")

#this file is not a CSV despite its csv file extension
red_wine_data <- read_delim("data/red-wine.csv", delim=";")
#total 12 columns 
colnames(red_wine_data) <- col_names_data$Variables

red_wine_data$type <- c("red")

combined_wine_tbl <- dplyr::bind_rows(white_wine_data, red_wine_data)

combined_wine_tbl |>
  filter(quality > 6.5 , alcohol < 13.2) |>
  arrange(desc(quality)) |>
  select(everything()) |>
  select(contains("acid"), alcohol, type, quality) |>
  group_by(quality) |>
  mutate( #keeps original columns 
    across(
      alcohol,
      list(mean = mean, median = median)
    )
  ) |>
  select(everything())

```

# Task 3: Database Practice

## Questions: 11, 12, & 13

```{r}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "data/lahman.db")
dbListTables(con)

tbl(con, "Teams")

tbl(con,sql(
  "
  SELECT * FROM `Teams`
  "
))



```

## Questions: 14

```{r}



tbl(con,sql(
  "
  SELECT h.playerID, yearID, category  FROM HallOfFame as h
  WHERE inducted = 'Y' 
  "
))



```

## Questions: 15

```{r}



tbl(con,sql(
  "
  SELECT p.playerID, `nameFirst`, `nameLast`, `yearID`, `category`  FROM `HallOfFame` as h
  INNER JOIN `People` as p
  ON h.playerID = p.playerID
  WHERE `inducted` = 'Y' 
  "
))



```

## Questions: 16

```{r}

tbl(con,sql(
  "
  SELECT m.playerID, sum(G) as G_managed , sum(W) as Total_W, sum(L) as Total_L,
         sum(W) * 1.0 /sum(L) as win_loss_ratio,  sum(W) * 100.0/sum(G) as career_win_loss_percentage
  FROM  Managers as m
  group by m.playerID
  order by win_loss_ratio desc 
  "
))




```

## Questions: 17

```{r}



tbl(con,sql(
  "
      SELECT h.playerID, p.nameFirst, p.nameLast, h.yearID,
             sum(G) as G_managed , sum(W) as Total_W, sum(L) as Total_L,
         sum(W) * 1.0 /sum(L) as win_loss_ratio,  sum(W) * 100.0/sum(G) as career_win_loss_percentage
      FROM HallOfFame as h
       INNER JOIN `People` as p
       ON h.playerID = p.playerID
       JOIN  Managers as m
      ON h.playerID =  m.playerID
      WHERE h.inducted = 'Y' 
      GROUP BY m.playerID
      order by win_loss_ratio desc 
  "
))

dbDisconnect(con)


```
