---
title: "awu3_HW4"
format: html
editor: visual
---

## Task 1: Conceptual Questions

**1.)** Rewrite the following nested function call using baseR's chaining operator:

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length \<
1.55), Species)

```{r}
library(tidyverse)
iris |> select(starts_with("Petal"), Species) |> filter(Petal.Length < 1.55) |> arrange(Species)



```

**2.)**What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

-   Long from data has values that are repeated in the first column but wide form does not have repeating values in the first column which makes wide form have less rows but more columns and vice versa for long form. Long form is preferred for statistical analysis

## Task 2: Chaining

### Glass Data

**3.)** Start a chain on your data object that overwrites the Type_of_glass variable using mutate()

-   Create a factor version of the variable that replaces the numbers with string labels (that is,
    replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the
    variable descriptions above to give meaningful values).

-   This could also be done by first using an ifelse() type call and then converting to a factor.
    Either way is fine

    ```{r}

    # import tidyverse package
    library(tidyverse)

    # read glass dataset and create df
    glass_df <- read_delim( "https://www4.stat.ncsu.edu/online/datasets/glass.data", delim=",", col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K","Ca","Ba","Fe", "Glass_type"))

    # change value based on glass value
    glass_df <- glass_df |> mutate(Glass_type = case_when(
      Glass_type == 1 ~"building_windows_float_processed",
        Glass_type == 2 ~ "building_windows_non_float_processed",
        Glass_type == 3 ~ "vehicle_windows_float_processed",
        Glass_type == 4 ~ "vehicle_windows_non_float_processed",
        Glass_type == 5 ~ "containers",
        Glass_type == 6 ~ "tableware",
        Glass_type == 7 ~ "headlamps"
      ))


    ```

**4.)** Continue your chain and keep only observations where the Fe variable is less than 0.2 and the Type of

Glass is either “tableware” or “headlamp”

```{r}

glass_df |> filter(Fe < 0.2, Glass_type %in% c("tableware", "headlamp"))


```

### Yeast data

**5.)**Start a chain that removes the seq_name and nuc columns

```{r}

# read yeast data and create df
yeast_df <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class"))

# remove seq_name and nuc columns
yeast_df <- yeast_df |> select(-seq_name, -nuc)

yeast_df
```

**6.)** Continue your chain to add columns corresponding to the mean and median of each numeric variable (mcg, gvh, alm, mit, erl, pox, and vac) at each class grouping (see the across() function as we did in the dplyr video!).

```{r}
# create mean and median cols for numeric variables
yeast_df |> group_by(class) |> mutate(
  across(where(is.numeric), list(mean = mean, median=median), .names = "{.col}_{.fn}"))
    
```

### Wine Data

```{r}
# import readxl package
library(readxl)

# read in white wine csv and create df
white_wine_df <- read_excel("../data/white-wine.xlsx")

# get col names from the second sheet in white wine dataset
wine_col_names <- read_xlsx("../data/white-wine.xlsx", sheet = excel_sheets("../data/white-wine.xlsx")[2])

wine_col_names

# set new col names from the second sheet
colnames(white_wine_df) <- wine_col_names$Variables

white_wine_df$type <- "white"

# read in red wine csv 
red_wine_df <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim= ";")

# set col names with col names from the csv
colnames(red_wine_df) <- wine_col_names$Variables
red_wine_df$type <- "red"

full_wine_df <- bind_rows(white_wine_df, red_wine_df)

full_wine_df
```

**7.)** Start a chain on your new combined data object to filter the data to only look at high-quality wines (quality \> 6.5) and wines that have a reasonable alcohol value (alcohol \< 132).

```{r}

# filter based on conditions
full_wine_df <- full_wine_df |> filter(quality > 6.5, alcohol < 13.2)
```

**8.)** Continue your chain to now sort the data from highest quality to lowest.

```{r}

# sory based on quality desc
full_wine_df <- full_wine_df |> arrange(desc(quality))

```

**9.)**Continue your chain to select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable.

```{r}
# select specific columns in the df
full_wine_df <- full_wine_df |> select(contains("acid"), alcohol, type, quality)
```

**10.)** Continue your chain to add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

```{r}

# add mean and sd to alcohol col
full_wine_df |> group_by(quality) |> mutate(mean_alcohol = mean(alcohol),sd_alcohol = sd(alcohol))
```

### Task 3: Database Practice

**11.)**Connect to the database and then look at all of the tables in the database.

```{r}
library(DBI)
# connect to the db
con <- dbConnect(RSQLite::SQLite(), "../data/lahman.db")
dbListTables(con)
```

**12.)** Use the tbl() function and dplyr to return all of the data from the Teams table for the year 2015.

```{r}

# pull teams table and filter
tbl(con, "Teams") |> filter(yearID == 2015)
```

**13.)** Repeat the above by using the sql() function within tbl() (here you have to write actual SQL!).

```{r}

# sql to do the same as above
tbl(con, sql("SELECT *
             FROM Teams
             WHERE (yearID = 2015)"))
```

**14.)** Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the HallOfFame table, the inducted variable is important here.

```{r}
# pull HOF table and filter
HOF_tbl <- tbl(con, "HallOfFame") |> filter(inducted == 'Y') |>select(playerID, yearID, category)
HOF_tbl
```

**15.)** Combine the table from the previous question with the People table in order to have the nameFirst and nameLast names added to what was returned above.

```{r}
# join HOF tbl with people tbl by playerID
HOF_full_tbl <- inner_join(tbl(con, "People") |> select(playerID, nameFirst, nameLast),
                           HOF_tbl,
                           by = join_by(playerID == playerID)
)
HOF_full_tbl


```

**16.)**Return only the playerID (manager ID in this case), G, W, and L columns from the Managers table. Use chaining to then:

-   Determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of
    fame managers.

-   Create a new variable that is the career win/loss percentage (Total_W/G_managed). (Note: I did this
    after collect() otherwise the column type has to be set. . . )

-    Sort the resulting data by the win/loss percentage variable (from largest to smallest).

```{r}

# get managers table and find win/loss and career win/loss (don't know if I should keep win/loss col)
manager_tbl <- tbl(con, "Managers") |> select(playerID, G, W, L) |> 
                     group_by(playerID) |> 
                     mutate(win_loss = sum(W) * 1.0 / sum(L))|> 
                     collect() |> mutate(career_win_loss = sum(W) * 1.0/sum(G)) |> 
                     arrange(desc(career_win_loss))
manager_tbl
```

**17.)**Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their category in
the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just
return all of the variables from both of the above two questions - 9 total columns)

```{r}
# join HOF and managers table
inner_join(manager_tbl, as.tibble(HOF_full_tbl), by = join_by(playerID == playerID))

```
