---
title: "HW4 Group7"
format: html
editor: visual
---

## Task 1: Conceptual Questions

**1.)** Rewrite the following nested function call using baseR's chaining operator:

```{r}
# Convert the iris data to a tibble and sorting the results by species

library(tidyverse)
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
                 1.55), Species)
```

**Solution:**

```{r}
library(tidyverse)
iris |> 
  select(starts_with("Petal"), Species) |> 
  filter(Petal.Length < 1.55) |> 
  arrange(Species)

```

**2.)**What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

**Solution:**

- Long Format: Each of the observation is a record with variables stored in their own columns.
- Wide Format: Multiple observations for the same subject maybe spread across multiple columns.
- Long Format is generally preferred for statistical analysis because it makes data easier to filter, group and analyze.

## Task 2: Chaining

### Glass Data

Note: Use chaining where possible!

Glass data In HW 3 you read in the glass data from <https://www4.stat.ncsu.edu/~online/datasets/glass.data.> Let’s reread that data (use your code from HW 3) that also names the columns. Recall the last variable being Type of Glass with values of – 1 building_windows_float_processed, – 2 build- ing_windows_non_float_processed,–3vehicle_windows_float_processed,–4vehicle_windows_non_float_processed (none in this database), – 5 containers, – 6 tableware, – 7 headlamps.

**3.)** Start a chain on your data object that overwrites the Type_of_glass variable using mutate()

- Create a factor version of the variable that replaces the numbers with string labels (that is, replace 1 with “building_windows_float_processed”, 2 with “building_win. . . ”, etc. - see the variable descriptions above to give meaningful values).

- This could also be done by first using an ifelse() type call and then converting to a factor. Either way is fine

  **Solution:**

  ```{r}
  library(tidyverse)

  # Read the glass data and add column names
  glass_df <- read_delim(
    "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
    delim = ",",
    col_names = c(
      "Id", "RI", "Na", "Mg", "Al", "Si",
      "K", "Ca", "Ba", "Fe", "Glass_type"
    )
  )

  # Replace the glass type numbers with factor labels
  glass_df <- glass_df |>
    mutate(
      Glass_type = factor(
        Glass_type,
        levels = 1:7,
        labels = c(
          "building_windows_float_processed",
          "building_windows_non_float_processed",
          "vehicle_windows_float_processed",
          "vehicle_windows_non_float_processed",
          "containers",
          "tableware",
          "headlamps"
        )
      )
    )

  glass_df
  ```

**4.)** Continue your chain and keep only observations where the Fe variable is less than 0.2 and the Type of

Glass is either “tableware” or “headlamp”

**Solution:**

```{r}

glass_df |>
  filter(
    Fe < 0.2,
    Glass_type %in% c("tableware", "headlamps")
  )

```

### Yeast data

**5.)** Start a chain that removes the seq_name and nuc columns

**Solution:**

```{r}

# read yeast data and create df
yeast_df <- read_table("https://www4.stat.ncsu.edu/online/datasets/yeast.data", col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class"))

# remove seq_name and nuc columns
yeast_df <- yeast_df |> select(-seq_name, -nuc)

# Display the updated data
yeast_df
```

**6.)** Continue your chain to add columns corresponding to the mean and median of each numeric variable (mcg, gvh, alm, mit, erl, pox, and vac) at each class grouping (see the across() function as we did in the dplyr video!).

**Solution:**

```{r}
# Group the yeast observations by class and add the mean and median of each numeric variable within each class
yeast_df |> group_by(class) |> mutate(
  across(where(is.numeric), list(mean = mean, median=median), .names = "{.col}_{.fn}"))
    
```

### Wine Data

```{r}
# import readxl package
library(readxl)

# read in white wine data and create df
white_wine_df <- read_excel("data/white-wine.xlsx")

# get col names from the second sheet in white wine dataset
wine_col_names <- read_xlsx("data/white-wine.xlsx", sheet = excel_sheets("data/white-wine.xlsx")[2])

wine_col_names

# set new col names from the second sheet
colnames(white_wine_df) <- wine_col_names$Variables

white_wine_df$type <- "white"

# read in red wine csv 
red_wine_df <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim= ";")

# set col names with col names from the csv
colnames(red_wine_df) <- wine_col_names$Variables
red_wine_df$type <- "red"

full_wine_df <- bind_rows(white_wine_df, red_wine_df)

full_wine_df
```

**7.)** Start a chain on your new combined data object to filter the data to only look at high-quality wines (quality \> 6.5) and wines that have a reasonable alcohol value (alcohol \< 132).

**Solution:**

```{r}

# filter based on conditions
full_wine_df <- full_wine_df |>
  filter(quality > 6.5, alcohol < 13.2)

full_wine_df
```

**8.)** Continue your chain to now sort the data from highest quality to lowest.

**Solution:**

```{r}
# Sort the wines from highest quality to lowest quality
full_wine_df <- full_wine_df |>
  arrange(desc(quality))

full_wine_df

```

**9.)**Continue your chain to select only the variables that contain acid, the alcohol variable, the type variable, and the quality variable.

**Solution:**

```{r}
# Select acid related variables along with alcohol, type, and quality
full_wine_df <- full_wine_df |>
  select(contains("acid"), alcohol, type, quality)

full_wine_df
```

**10.)** Continue your chain to add the mean and standard deviation of the alcohol variable to the data set for each setting of the quality variable.

**Solution:**

```{r}

# Group the wines by quality and add the mean and standard deviation of alcohol for each quality group

full_wine_df |>
  group_by(quality) |>
  mutate(
    mean_alcohol = mean(alcohol),
    sd_alcohol = sd(alcohol)
  )
```

### Task 3: Database Practice

**11.)**Connect to the database and then look at all of the tables in the database.

**Solution:**

```{r}
library(DBI)
# Connect to the Lahman SQLite database
con <- dbConnect(RSQLite::SQLite(), "data/lahman.db")
dbListTables(con)
```

**12.)** Use the tbl() function and dplyr to return all of the data from the Teams table for the year 2015.

**Solution:**

```{r}
# Access the Teams table and keep only records from 2015

tbl(con, "Teams") |> filter(yearID == 2015)
```

**13.)** Repeat the above by using the sql() function within tbl() (here you have to write actual SQL!).

**Solution:**

```{r}
# Use SQL to return all Teams records from 2015

tbl(con, sql("SELECT *
             FROM Teams
             WHERE (yearID = 2015)"))
```

**14.)** Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their category (only those three variables!). See the HallOfFame table, the inducted variable is important here.

**Solution:**

```{r}
# Return hall of fame with the requested variables
HOF_tbl <- tbl(con, "HallOfFame") |>
  filter(inducted == "Y") |>
  select(playerID, yearID, category)

HOF_tbl
```

**15.)** Combine the table from the previous question with the People table in order to have the nameFirst and nameLast names added to what was returned above.

**Solution:**

```{r}
# Select player names from the People table and join them with the Hall of Fame results using playerID

HOF_full_tbl <- inner_join(
  tbl(con, "People") |>
    select(playerID, nameFirst, nameLast),
  HOF_tbl,
  by = join_by(playerID == playerID)
)

HOF_full_tbl


```

**16.)**Return only the playerID (manager ID in this case), G, W, and L columns from the Managers table. Use chaining to then:

- Determine the overall win/loss records (sum of wins and sum of losses) for each of these hall of fame managers.

- Create a new variable that is the career win/loss percentage (Total_W/G_managed). (Note: I did this after collect() otherwise the column type has to be set. . . )

- Sort the resulting data by the win/loss percentage variable (from largest to smallest).

  **Solution:**

```{r}

# Select the manager ID, games, wins, and losses and calculate career totals for each manager

manager_tbl <- tbl(con, "Managers") |>
  select(playerID, G, W, L) |>
  group_by(playerID) |>
  summarize(
    G_managed = sum(G, na.rm = TRUE),
    Total_W = sum(W, na.rm = TRUE),
    Total_L = sum(L, na.rm = TRUE)
  ) |>
  collect() |>
  
  mutate(career_win_loss = Total_W / G_managed) |>
  
# Sort from highest to lowest career win/loss percentage
  arrange(desc(career_win_loss))

manager_tbl
```

**17.)**Use the results from the previous two questions to answer this one! Return information for only the people that managed a team and were inducted into the hall of fame (regardless of their category in the hall of fame - you want to return all people in the hall of fame that ever managed at all). (Just return all of the variables from both of the above two questions - 9 total columns)

**Solution:**

```{r}
# Join HOF and managers table
hof_managers <- HOF_full_tbl |>
  collect() |>
  inner_join(
    manager_tbl,
    by = "playerID"
  )
hof_managers
```
