---
title: "hw2_final"
format: html
editor: visual
---

## **Task 1: Basic Vector Practice**

**Q1)** Create two vectors. One vector corresponding to the pre measurements and one to the post measurements.

```{r}

# Pre treatment blood pressure measurements
pre_bp <- c(130, 128, 116, 124, 133, 
            134, 118, 126, 114, 127, 
            141, 138, 128, 140, 137, 
            131, 120, 128, 139, 135)

# Post-treatment blood pressure measurements
post_bp <- c(114, 98, 113, 99, 107, 116,
             113, 111, 119, 117, 101, 119, 
             130, 122, 106, 106, 124, 102, 
             117, 113)

```

**Q2)** Assign names to the vector elements using the paste() function.

```{r}

# create names and asign them
patient <- paste('Subject', 1:20, sep = '_')

names(pre_bp) <- patient
names(post_bp) <- patient

pre_bp
post_bp

```

**Q3)** Calculate the change in blood pressure for each patient by subtracting post-treatment measurements from pre-treatment measurements

```{r}
# Calculate the change in blood pressure 
diff_bp <- pre_bp - post_bp
diff_bp
```

**Q4)** Calculate the average decrease in blood pressure across all patients

```{r}

# Calculate avg change for all patients
avg_dec <- mean(diff_bp)
avg_dec
```

**Q5)** Determine which patients experienced a decrease in blood pressure after treatment (a positive change). Use the which() function to just return the indices (and names) associated with this type of change.

```{r}

# Find patients with a positive change
which(diff_bp > 0)
```

**Q6)** Subset the vector of differences to only return those that have a positive change.

```{r}

# Subset to keep only positive changes
pos_change <- diff_bp[diff_bp > 0]

pos_change
```

**Q7)** Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive change).

```{r}

# Find the mean positive change
avg_pos_change <- mean(pos_change)
avg_pos_change
```

------------------------------------------------------------------------

## **Task 2: Basic Data Frame Practice**

**1)** Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp

```{r}

# Create bp dataframe
bp_df <- data.frame(
  patient = patient,
  pre_bp = pre_bp,
  post_bp = post_bp,
  diff_bp = diff_bp
)

bp_df
```

**2)** Return only rows where the diff_bp column is negative. (Use \[ or learn about the subset() function if you’d like.

```{r}

# Return rows where bp change is negative
bp_df[bp_df$diff_bp < 0, ]
```

**3)** Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120. Recall you can use \$ to reference a column.

```{r}
# Create a new column for optimal bp
bp_df$healthy_bp <- bp_df$post_bp < 120

bp_df
```

**4)** Finally, print the data frame out nicely in your final document by modifying the code below appropriately.

```{r}

#print df formatted

knitr::kable(bp_df)
```

------------------------------------------------------------------------

## **Task 3: List Practice**

**1)** Create a new data frame with this data in it that is similar to the data frame from task 2

```{r}

# pre-treatment placebo bp measurements
pre_placebo_bp <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

#post treatment placebo bp measurements
post_placebo_bp <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

#calculate the change in bp
placebo_diff <- pre_placebo_bp - post_placebo_bp

#get the subject names
placebo_names <- paste("Subject", 1:10, sep="_")

# create placebo df
placebo_df <- data.frame(
  patient = placebo_names,
  pre_bp = pre_placebo_bp,
  post_bp = post_placebo_bp,
  diff_bp = placebo_diff
)

#add column for optimal bp 
placebo_df$healthy_bp <- placebo_df$post_bp <120

placebo_df
```

**2)** Now create and store a list with two elements:

-   1st element named treatment and contains the first data frame you created.

-   2nd element named placebo and contains the second data frame you created

```{r}

# Create list with treatment and placebo df
bp_list <- list(treatment = bp_df,
                placebo = placebo_df)
bp_list
```

**3)** Access the first list element using three different types of syntax

```{r}
# access using df literal name
bp_list$treatment

# access using df position
bp_list[[1]]

# access using df name
bp_list[["treatment"]]
```

**4)** In one line, access the placebo data frame, pre_bp column

```{r}
# access pre_bp column in placebo df
bp_list$placebo$pre_bp
```

------------------------------------------------------------------------

## **Task 4: Control Flow Practice**

**1)** Suppose we want to characterize the post-treatment (or placebo) blood pressure measurement as optimal (≤ 120), borderline (120 \< bp ≤ 130), and high (\> 130). First, create a new column in each data frame from above called status.

```{r}

# Create new empty column 'status' for treatment and placebo data
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)

bp_list$treatment
bp_list$placebo
```

**2)** For the non-placebo data frame (within the list), create a for loop and use if/then/else logic to create the status column’s values.

```{r}

# assign values to the status column by classifying bp for treatment df
for (i in 1:20) {
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  }
  else if (bp_list$treatment$post_bp[i] <= 130) {
    bp_list$treatment$status[i] <- "borderline"
  } else {
    bp_list$treatment$status[i] <- "high"
  }
}
bp_list$treatment

```

**3)**Repeat for the placebo data frame (within the list)

```{r}

# assign values to the status column by classifying bp for placebo df
for (i in 1:10) {
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  }
  else if (bp_list$placebo$post_bp[i] <= 130) {
    bp_list$placebo$status[i] <- "borderline"
  } else {
    bp_list$placebo$status[i] <- "high"
  }
}
bp_list$placebo
```
