---
title: "hw5_group7_final"
format: html
editor: visual
---

# TASK ONE: QUERYING AN API AND A TIDY-STYLE FUNCTION

*Question 1*

Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

```{r}
library(httr)
library(tidyverse)
library(dotenv)

# load in env file for API key
load_dot_env()

api_key <- Sys.getenv("NEWS_API_KEY")


articles <- GET(
  "https://newsapi.org/v2/everything",
  query = list(
    q = 'biology',
    from = '2026-09-01',
    language = 'en',
    pagesize = 50,
    apiKey = api_key
  )
)

```

*Question 2*

Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
articles_content <- content(
  articles,
  as = 'parsed',
  simplifyVector = FALSE
)

names(articles_content)

article_info <- articles_content$articles

article_info[[1]]
```

```{r}
article_tibble <- tibble(
  source = map(article_info, 'source'),
  author = map_chr(article_info, 'author', .default = NA_character_),
  title = map_chr(article_info, 'title', .default = NA_character_),
  description = map_chr(article_info, 'description', .default = NA_character_),
  publishedAt = map_chr(article_info, 'publishedAt', .default = NA_character_)
)

str(article_tibble)
```

*Question 3*

Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
news_api_search <- function(subject, from_date, api_key){
  API_article <- GET(
    "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = from_date,
      apiKey = api_key
    )
  )
  article_response <- content(
    API_article,
    as = 'parsed',
    simplifyVector = FALSE
  )
  article_response$articles
}

physics_article <- news_api_search('physics', '2026-09-01', api_key)
physics_article[[1]]

epidemiology_article <- news_api_search('physics', '2026-09-01', api_key)
epidemiology_article[[1]]
```

# TASK TWO: SUMMARIZING ROLLER COASTER DATA

*Reading in data and changing variables*

```{r}
library(readr)
rollercoaster_data <-  read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")
rollercoaster_data
```

```{r}
library(dplyr)

rollercoaster_data <- rollercoaster_data |>
  mutate(
    StateF = as.factor(State),
    Census_RegionF = as.factor(`Census Region`),
    AgeGroupF = as.factor(AgeGroup),
    ScaleF = as.factor(Scale),
    TypeF = as.factor(Type),
    DesignF = as.factor(Design),
    MakeF = as.factor(Make)
  )
rollercoaster_data
```

## BASIC EDA

*Question 4*

Look at how the data is stored and see if everything makes sense.

```{r}
glimpse(rollercoaster_data)
```

*Question 5*

Document any missing values in the data

```{r}
na_total <- function(column){
  sum(is.na(column))
}

rollercoaster_na <- rollercoaster_data |>
  summarize(across(everything(), na_total))
rollercoaster_na

 missing_values <- colSums(is.na(rollercoaster_data))

    # Display the missing value counts
    missing_values
```

## CATEGORICAL VARIABLES

*Question 6*

- Interpret a number from each resulting table( that is pick out a value produced and explain what that value means.)

- **1 Way Contingency Table**-\> There are 69 roller coasters in the older age group the dataset

  **2 Way Contingency Table** -\> There are 36 roller coasters in the older age group that are made of steel

  **3 Way Contingency Table** -\> There are 79 new roller coasters made of steel meant for a thrill ride

  ```{r}

  #one way contigency table
  table(rollercoaster_data$AgeGroupF)

  #two way contigency table
  table(rollercoaster_data$AgeGroupF, rollercoaster_data$TypeF)

  #three way contigency table
  table(rollercoaster_data$AgeGroupF, 
        rollercoaster_data$TypeF, 
        rollercoaster_data$ScaleF)
  ```

*Question 7*

Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods: • Once, by subsetting the data (say with filter()) and then creating the two-way table • Once, by creating a three-way table and subsetting it

- Once, by subsetting the data and then creating the two-way table
- Once, by creating a three-way table and subsetting it

```{r}
# subset data then create two-way table
new_coaster <- rollercoaster_data |> filter(AgeGroupF == "3:Newest")

# new coasters split up by region and scale
table(new_coaster$Census_RegionF, new_coaster$ScaleF)

#three-way table and subset
three_way <- table(rollercoaster_data$Census_RegionF, 
                   rollercoaster_data$ScaleF, 
                   rollercoaster_data$AgeGroupF)
three_way[,,"3:Newest"]
```

*Question 8*

Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result look more like the output from table()

```{r}
  # Count coasters for each combination of AgeGroup and Inversions
  age_inversion_counts <- rollercoaster_data |>
    group_by(AgeGroup, `Inversions?`) |>
    summarize(Count = n(), .groups = "drop")

  age_inversion_counts
```

```{r}
    # Reshape the counts to look like a contingency table
    age_inversion_wide <- age_inversion_counts |>
      pivot_wider(
        names_from = `Inversions?`,
        values_from = Count
      )

    age_inversion_wide
```

*Question 9*

Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots

```{r}
library(ggplot2)

# Create a stacked bar graph
ggplot(data = rollercoaster_data, 
            aes(x = Census_RegionF, fill = ScaleF)) + 
  geom_bar() +
  labs(x = "Census Region", 
       y = "Count", 
       title = "Type of Coaster Count in Each Region") 


# Create a side by side bar graph
ggplot(data = rollercoaster_data, aes(x = Census_RegionF, fill = ScaleF)) + 
  geom_bar(position = "dodge") +
  labs(x = "Census Region", 
       y = "Count", 
       title = 'Rollercoaster Statistics: Design and Type', x = 'Design', y = '# of Rollercoasters', fill = 'Type')
```
