---
title: "awu3_HW5"
format: html
editor: visual
---

## Task 1: Querying an API and a Tidy-Style Function

**1.)**Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the
past with a free account.

```{r}
library(httr)
library(dotenv)

# load in env file for API key
load_dot_env()

api_key <- Sys.getenv("API_KEY")

#url for get
url <- paste0("https://newsapi.org/v2/everything?q=Iran&from=2026-08-22&language=en&pageSize=50&apiKey=", api_key)

# get data and check structure
info <- httr::GET(url)
str(info, max.level = 1)

```

**2.)**Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
library(jsonlite)
library(dplyr)

# parse json and convert to char
parsed <- fromJSON(rawToChar(info$content))
parsed

#convert to tibble
news <- as_tibble(parsed$articles)
news

```

**3.)**Now write a quick function that allows the user to easily query this API. The inputs to the function
should be the title/subject to search for (string), a time period to search from (string - you’ll search
from that time until the present), and an API key.

Use your function twice to show it works!

```{r}
library(dplyr)
#function to query API
get_news <- function(subject, from_time, key) {
  info <- GET(
    "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = from_time,
      apiKey = key
    )
  )
  parsed <- fromJSON(rawToChar(info$content))
  parsed
  
  as_tibble(parsed$articles)
}
```

```{r}

#get us news
get_news("United States", "2026-08-30", api_key)

#get Apple news
get_news("Apple", "2026-08-25", api_key)
```

## Task 2 - Summarizing Roller Coaster Data

#### Basic EDA

**4.)** Look at how the data is stored and see if everything makes sense

```{r}
library(readr)

#read in coaster data
coaster_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

# convert some cols to factor columns
coaster_data <- coaster_data |> mutate(
  State = as.factor(State),
  `Census Region` = as.factor(`Census Region`),
  AgeGroup = as.factor(AgeGroup),
  Scale = as.factor(Scale),
  Type = as.factor(Type),
  Design = as.factor(Design),
  Make = as.factor(Make)
)
coaster_data
```

**5.**) Document any missing values in the data

```{r}
#helper function to count missing values
sum_na <- function(column){
  sum(is.na(column))
}

#count missing values for each column
coaster_data |> summarise(across(everything(), sum_na))

```

#### Categorical Variables

**6.)** Create a one-way contingency table, a two-way contigency table, and a three-way contingency table for some of the factor variables you created previously. Use table() and accomplish this

-   Interpret a number from each resulting table( that is pick out a value produced and explain what that value means.)

-   **1 Way Contingency Table**-\> There are 69 roller coasters in the older age group the dataset

    **2 Way Contingency Table** -\> There are 36 roller coasters in the older age group that are made of steel

    **3 Way Contingency Table** -\> There are 79 new roller coasters made of steel meant for a thrill ride

    ```{r}

    #one way contigency table
    table(coaster_data$AgeGroup)

    #two way contigency table
    table(coaster_data$AgeGroup, coaster_data$Type)

    #three way contigency table
    table(coaster_data$AgeGroup, 
          coaster_data$Type, 
          coaster_data$Scale)
    ```

**7.)**Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods:

-   Once, by subsetting the data and then creating the two-way table

-   Once, by creating a three-way table and subsetting it

```{r}
# subset data then create two-way table
new_coaster <- coaster_data |> filter(AgeGroup == "3:Newest")

# new coasters split up by region and scale
table(new_coaster$`Census Region`, new_coaster$Scale)

#three-way table and subset
three_way <- table(new_coaster$`Census Region`, 
                   new_coaster$Scale, 
                   new_coaster$AgeGroup)
three_way[,,"3:Newest"]
```

**8.)**Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result look more like the output from table().

```{r}
library(tidyr)
#2 way table by grouping and summaerize
coaster_data |> 
  group_by(Scale, `Census Region`) |>
  summarize(count = n()) |>
  pivot_wider(names_from = Scale, values_from = count)
```

**9.)** Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for
the plots.

```{r}
library(ggplot2)

# Create a stacked bar graph
g <- ggplot(data = coaster_data, 
            aes(x = `Census Region`, fill = Scale)) + 
  geom_bar() +
  labs(x = "Census Region", 
       y = "Count", 
       title = "Type of Coaster Count in Each Region") 
g

# Create a side by side bar graph
g2 <-ggplot(data = coaster_data, aes(x = `Census Region`, fill = Scale)) + 
  geom_bar(position = "dodge") +
  labs(x = "Census Region", 
       y = "Count", 
       title = "Type of Coaster Count in Each Region") 
g2
```
