---
title: "ST558_HW5"
author: "Lasya Harini Banka"
format: html
editor: visual
---

## TASK1:

1.  Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

    **Solution:**

    ```{r}
    library(tidyverse)
    library(httr)
    my_key <- Sys.getenv("NEWS_API_KEY")

    # Request recent technology news
    news <- GET(
      "https://newsapi.org/v2/everything",
      query = list(
        q = "technology",
        from = "2026-09-01",
        language = "en",
        pageSize = 50,
        apiKey = my_key
      )
    )

    # Check if the request was successful
    status_code(news)

    ```

    I searched NewsAPI for English-language articles about technology published since September 1, 2026. The response from the API was stored in news.

2.  Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

    **Solution:**

    ```{r}
    # Parse the API response
    news_content <- content(
      news,
      as = "parsed",
      simplifyVector = FALSE
    )

    # Look at the information returned
    names(news_content)
    # Get the article information
    articles <- news_content$articles

    # Look at the first article
    articles[[1]]
    ```

    ```{r}
    # Convert the article information into a tibble
    article_df <- tibble(
      source = map(articles, "source"),
      author = map_chr(articles, "author", .default = NA_character_),
      title = map_chr(articles, "title", .default = NA_character_),
      description = map_chr(articles, "description", .default = NA_character_),
      publishedAt = map_chr(articles, "publishedAt", .default = NA_character_)
    )

    # Check the structure
    glimpse(article_df)
    ```

    I parsed the API response and extracted the article information into a tibble. The first column, source is a list column containing information about the source of each article.

3.  Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

    **Solution:**

    ```{r}
    # Create a function to search NewsAPI
    search_news <- function(subject, from_date, api_key) {
      
      # Send a request using the inputs provided
      response <- GET(
        "https://newsapi.org/v2/everything",
        query = list(
          q = subject,
          from = from_date,
          language = "en",
          pageSize = 50,
          apiKey = api_key
        )
      )
      
      # Stop if the API request fails
      stop_for_status(response)
      
      # Parse the response
      result <- content(
        response,
        as = "parsed",
        simplifyVector = FALSE
      )
      
      # Return the articles
      result$articles
    }
    ```

    ```{r}
    # Search for data science news
    data_science_news <- search_news(
      subject = "data science",
      from_date = "2026-09-01",
      api_key = my_key
    )

    # Look at the first result
    data_science_news[[1]]
    ```

    ```{r}
    # Search for space news
    space_news <- search_news(
      subject = "space",
      from_date = "2026-09-01",
      api_key = my_key
    )

    # Look at the first result
    space_news[[1]]
    ```

## TASK2:

4.  Look at how the data is stored and see if everything makes sense.

    **Solution:**

    ```{r}
    coasters <- read_csv(
      "https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv",
      show_col_types = FALSE
    )
    # Convert categorical variables to factors
    coasters <- coasters |>
      mutate(
        State = as.factor(State),
        `Census Region` = as.factor(`Census Region`),
        AgeGroup = as.factor(AgeGroup),
        Scale = as.factor(Scale),
        Type = as.factor(Type),
        Design = as.factor(Design),
        `Inversions?` = as.factor(`Inversions?`),
        Make = as.factor(Make)
      )

    # Check the structure after converting the variables
    glimpse(coasters)
    ```

    The dataset has 635 roller coasters and 21 variables. I converted the categorical variables to factors, while boundary remains a character variable because it contains geographic information.

5.  Document any missing values in the data.

    **Solution:**

    ```{r}
    missing_values <- colSums(is.na(coasters))

    # Display the missing value counts
    missing_values
    ```

    Missing values were found in TopSpeed (74), MaxHeight(8), Drop (128), Length (33), Duration (66), and Make (36). The remaining variables have no missing values.

6.  Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this. • Interpret a number from each resulting table (that is, pick out a value produced and explain what that value means.)

    **Solution:**

    ```{r}
    # One-way contingency table for age group
    age_table <- table(coasters$AgeGroup)

    age_table
    ```

    The one-way table shows that 376 roller coasters belong to the 3:Newest age group.

    ```{r}
    # Two-way contingency table for age group and inversions
    age_inversion_table <- table(
      coasters$AgeGroup,
      coasters$`Inversions?`
    )

    age_inversion_table
    ```

    From the two-way table, 70 roller coasters in the 2:Recent age group have inversions.

    ```{r}
    # Three-way contingency table
    three_way_table <- table(
      coasters$AgeGroup,
      coasters$`Inversions?`,
      coasters$`Census Region`
    )

    three_way_table
    ```

    From the three-way table, 30 roller coasters in the 2:Recent age group in the South have inversions.

7.  Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods: • Once, by subsetting the data (say with filter()) and then creating the two-way table • Once, by creating a three-way table and subsetting it

    **Solution:**

    ```{r}
    # Method 1: Filter the data to the South region
    south_coasters <- coasters |>
      filter(`Census Region` == "South")

    # Create the two-way table
    south_table1 <- table(
      south_coasters$AgeGroup,
      south_coasters$`Inversions?`
    )

    south_table1
    ```

    ```{r}
    # Method 2: Select the South region from the three-way table
    south_table2 <- three_way_table[, , "South"]

    south_table2
    ```

    Both methods produce the same conditional two-way table for roller coasters in the South.

8.  Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result look more like the output from table().

    **Solution:**

    ```{r}
    # Count coasters for each combination of AgeGroup and Inversions
    age_inversion_counts <- coasters |>
      group_by(AgeGroup, `Inversions?`) |>
      summarize(Count = n(), .groups = "drop")

    age_inversion_counts
    ```

    ```{r}
    # Reshape the counts to look like a contingency table
    age_inversion_wide <- age_inversion_counts |>
      pivot_wider(
        names_from = `Inversions?`,
        values_from = Count
      )

    age_inversion_wide
    ```

9.  Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.

    **Solution:**

    ```{r}
    # Stacked bar graph
    ggplot(coasters, aes(x = AgeGroup, fill = `Inversions?`)) +
      geom_bar() +
      labs(
        title = "Roller Coasters by Age Group and Inversion Status",
        x = "Age Group",
        y = "Number of Roller Coasters",
        fill = "Has Inversions?"
      )
    ```

    ```{r}
    # Side-by-side bar graph
    ggplot(coasters, aes(x = AgeGroup, fill = `Inversions?`)) +
      geom_bar(position = "dodge") +
      labs(
        title = "Roller Coasters by Age Group and Inversion Status",
        x = "Age Group",
        y = "Number of Roller Coasters",
        fill = "Has Inversions?"
      )
    ```
