---
title: "hw5_bdevapa"
format: html
---

# Task 1 - Querying an API and a Tidy-Style Function

## Question 1

Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account

```{r}
#1
library(httr)
resp <- GET("https://newsapi.org/v2/everything?q=ai&from=2026-09-01&language=en&pageSize=50&apiKey=43416691f7e147da9122be62d78cff7c")

```

## Question 2

Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
#2

library(jsonlite)
library(tidyverse)
resp_json <- fromJSON(rawToChar(resp$content))
news_articles <- as_tibble(resp_json$articles)
news_articles

```

## Question 3

Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
#3

find_news_articles <- function(topicStr, fromDateStr, apiKey) {
  
  #check for missing inputs 
  if (missing(topicStr) || missing(fromDateStr) || missing(apiKey)) {
    stop("topicStr, fromDateStr, and apiKey are required")
  }
  
   # Convert the supplied date to a Date -- without supplying format the default format is not what not acceptable by the api
  fromDate <- as.Date(fromDateStr, format = "%m-%d-%Y")

  # Check that date is not more than 30 days old
  #wihtout the error it simply gives 0 results -- the api doesnbt provide results for last 30 days
  if (fromDate < Sys.Date() - 30) {
    stop("fromDateStr cannot be more than 30 days old")
  }
  
  library(httr)
  library(jsonlite)
  
  url <- paste0(
    "https://newsapi.org/v2/everything?q=", topicStr,
    "&from=", fromDateStr,
    "&language=en&pageSize=50&apiKey=", apiKey
    )
  resp <- GET(url)
 
resp_json <- fromJSON(rawToChar(resp$content))
news_articles <- as_tibble(resp_json$articles)

}

news <- find_news_articles("tesla","09-01-2026","43416691f7e147da9122be62d78cff7c")
news

```

# Task 2 - Summarizing Roller Coaster Data

- Basic EDA

## Question 4 & 5
- Look at how the data is stored and see if everything makes sense.
- Document any missing values in the data.


```{r}
#4
 library(readr)
 roller_coaster_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")
 
str(roller_coaster_data)



# the data generally makes sense, some columns such as yearOpened can be character
roller_coaster_data <- roller_coaster_data |>
  mutate(YearOpened = as.character(`Year Opened`),
         StateF = factor(State),
         AgeGroupF = factor(AgeGroup),
         ScaleF = factor(Scale),
         TypeF = factor(Type),
         DesignF = factor(Design)
         )

roller_coaster_data

# 5 - document missing values 

sum_na <- function(column){
  sum(is.na(column))
}

na_counts <- roller_coaster_data |>
  summarize(across(everything(), sum_na))
na_counts


```

- Categorical variables

## Question - 6

Create a one-way contingency table, a two-way contingency table, and a three-way contingency table
for some of the factor variables you created previously. Use `table()` to accomplish this

```{r}
#6
# group by type factor variable and counting for each group
one_way <- table(roller_coaster_data$TypeF) 

one_way

# two way table -- group by roller coaster and by age group factor variables
two_way <- table(roller_coaster_data$AgeGroupF, roller_coaster_data$TypeF)

two_way

# three way table -- group by roller caoster and by agen group factor variables -- but then also group by the design 
three_way <- table(roller_coaster_data$AgeGroupF, roller_coaster_data$TypeF, roller_coaster_data$DesignF)

three_way


```

## Question - 7

- Create a conditional two-way table using table(). That is, condition on one variable’s setting and
create a two-way table. Do this using two different methods:
• Once, by subsetting the data (say with filter()) and then creating the two-way table
• Once, by creating a three-way table and subsetting it

```{r}
#7
# by subsetting the data (say with filter()) and then creating the two-way table
bobsled_data <- roller_coaster_data |>
  filter(DesignF == "Bobsled")

conditional_two_way_7_1 <- table(bobsled_data$AgeGroupF, bobsled_data$TypeF)

conditional_two_way_7_1

# by creating a three-way table and subsetting it
three_way <- table(roller_coaster_data$AgeGroupF, roller_coaster_data$TypeF, roller_coaster_data$DesignF)


conditional_two_way_7_2 <- three_way[, , "Bobsled"]

conditional_two_way_7_2 



```
## Question - 8

Create a two-way contingency table using group_by() and summarize() from dplyr. Then use
pivot_wider() to make the result look more like the output from table().

```{r}

#8

two_way <- table(roller_coaster_data$AgeGroupF, roller_coaster_data$TypeF)

two_way

two_way_tidy <- roller_coaster_data |>
  group_by(AgeGroupF, TypeF) |>
  summarize(count = n()) |>
  pivot_wider(names_from = TypeF, values_from = count)


two_way_tidy


```


## Question - 9

Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for
the plots.

```{r}

#9




g <- ggplot(roller_coaster_data, aes(x = AgeGroupF, fill = TypeF))
g + geom_bar() + 
  labs(title = "Roller Coaster Type by Age Group") +
  scale_fill_discrete("Type of material")

#Below is antoer way to do it 

g1 <- ggplot(data = roller_coaster_data |> 
         group_by(AgeGroupF, TypeF) |>
         summarize(count = n()), aes(x = AgeGroupF, y = count, fill = TypeF)) 
g1 + geom_bar(stat = "identity") +
  labs( title = "Roller Coaster Type by Age Group (using Stat Layer)",
        x = "Age Group") +
  scale_fill_discrete("Type of material")


```